\input{common/preamble.tex}
\color{dark}
\setlength{\parindent}{0em}
\setlength{\parskip}{0.5em}
\renewcommand{\baselinestretch}{1.2}

\begin{document}
\centering
\LARGE
Arch Linux installation guide for MacBook Pro Retina\\
 - mid-2015 model - 
\normalsize\justify
\tableofcontents
\clearpage

\section{Forewords}

\noindent After lots of reading, searching, experimenting, furious late-night shell command typing and 
do-overs here are the results of replacing OSX with Arch Linux (KDE) on a mid-2015 MacBook Pro. 

\noindent This guide is a record of my experience and results. I'm leaving this out there for anyone that
might find this useful.

\noindent I would also advise backing up your drive using a complete bit-to-bit cloning process (unlike me...oops) so that you retain a copy of everything including the recovery partition on the Mac. 
It's better to be safe...

\section{Machine Specs}

\begin{tabularx}{\textwidth}{|c|X|}
	\hline
	Display   & 15.4" LED-backlit Retina display with IPS technology; 2880-by-1800 native resolution at 220 pixels per inch with support for millions of colours \\\hline
	Processor & 2.5GHz quad-core Intel Core i7 processor (Turbo Boost up to 3.7GHz) with 6MB shared L3 cache \\\hline
	RAM       & 16GB of 1600MHz DDR3L memory \\\hline
	GPU       & Intel Iris Pro Graphics, <br>AMD Radeon R9 M370X with 2GB of GDDR5 memory \\\hline
	Storage   & 512GB PCIe-based flash \\\hline
	Webcam    & 720p FaceTime HD camera \\\hline
	Network   & 802.11ac Wi‑Fi wireless networking; IEEE 802.11a/b/g/n compatible, \linebreak
				Bluetooth 4.0 wireless technology \\\hline
\end{tabularx}

\section{Results}

//TODO

//Heating issues with mbp, fan control and turning off boost on the processor to keep things
reasonable temperature-wise.

\section{Pre-Installation}

\subsection{Preparations}

\begin{itemize-steps}
	\item \textbf{Backup drive} (complete drive clone preferable).
	\item Make a copy of the colour profile file(s) on the mac located in \code{/Library/ColorSync/Profiles/Displays/*} to a USB stick. It will be useful later on Linux.

\end{itemize-steps}

\subsection{Making a bootable USB}

\subsubsection{From Linux}

\begin{itemize-steps}
	\item \terminalcmd{dd if=archlinux.iso of=/dev/sdX bs=16M \&\& sync} \\
	      \block{Where \code{X} is your target USB drive letter (use \code{lsblk} for an overview of all connected drives to find out)}
\end{itemize-steps}

\subsubsection{From a Mac}

\begin{itemize-steps}
	\item \terminalcmd{diskutil unmountDisk diskX} \\
		  \block{Where \code{X} is your target USB drive number (use \code{diskutil list} for an overview of all connected drives to find out).}
	\item \terminalcmd{sudo dd if=/Users/\$USERNAME/Downloads/archlinux.iso of=/dev/diskX bs=16M}\\
		  \block{Replace \code{\$USERNAME} with your username on the mac and replace \code{X} with the USB drive's number.}
\end{itemize-steps}

\section{Installation}

\subsection{Booting from the USB stick}

Simply plug in the USB in the MBP and press the \fbox{Alt} (a.k.a. \fbox{options}) key during start up
to reach the boot menu.

\subsection{Preliminary setup}

\begin{itemize-steps}
	\item \terminalcmd{loadkeys uk}\\
		  \block{Load your keyboard layout. Replace `uk` with whichever you have on your machine.}
	\item \terminalcmd{wifi-menu}\\
		  \block{Connect up to the wifi network. On the MPB Pro 2015 the wifi gets detected and works out of the box at this stage.}
\end{itemize-steps}

\subsection{Partitioning and Crypto volume setup}

The assumption here is that the hard drive where everything will be installed to is \code{/dev/sda}. The partition structure will thus be: \\
\code{/dev/sda1} for the EFI partition,\\
\code{/dev/sda2} for the Boot partition and\\
\code{/dev/sda3} for the encrypted volume where the root, home and swap will be located.\\

\begin{itemize-steps}
	\item \terminalcmd{cgdisk /dev/sda}
	      \begin{blocksection}
		  \blockheader{Partition layout with just Linux}
		  \textit{Warning: this will erase every thing including the recovery partition for OSX. \\
		  No easy way to go back after this without a bit-to-bit HDD clone}
		  \begin{enumerate}
		  	\item 100M partition `EFI' (Hex \#\code{ef00})
		  	\item 250M partition `Boot' (Hex \#\code{8300})
		  	\item 100\% remainder of the space for the crypto volume (Hex \#\code{8300})
		  \end{enumerate}
		  The root and home partition will be created later in the crypto volume
	  	\end{blocksection}
	\item \terminalcmd{cryptsetup --verbose --verify-passphrase --cipher aes-xts-plain64 --key-size 512 --hash sha512 
		--iter-time 5000 --use-random luksFormat /dev/sda3}
		  \block{Sets up the crypto volume.}
	\item \terminalcmd{cryptsetup open --type luks /dev/sda2 mctoasty}
		  \block{\code{mctoasty} is the mounted name of the crypto volume. You can choose your own.}
\end{itemize-steps}

\subsubsection{Create crypto volume partitions}
\begin{itemize-steps}
	\item \terminalcmd{pvcreate /dev/mapper/mctoasty}
	\item \terminalcmd{vgcreate vg0 /dev/mapper/mctoasty}
	\item \terminalcmd{lvcreate --size 16G vg0 --name swap}
	\item \terminalcmd{lvcreate --size 50GB vg0 --name root}
		  \block{Adjust root's size as required.}
	\item \terminalcmd{lvcreate -l +100\%FREE vg0 --name home}
\end{itemize-steps}

\subsubsection{Formatting all the partitions}

\begin{itemize-steps}
	\item \terminalcmd{mkfs.vfat -F32 /dev/sda1}
		  \block{EFI partition}
	\item \terminalcmd{mkfs.ext2 /dev/sda2}
		  \block{Boot partition}
	\item \terminalcmd{mkswap /dev/mapper/vg0-swap}
	\item \terminalcmd{mkfs.ext4 /dev/mapper/vg0-root}
	\item \terminalcmd{mkfs.ext4 /dev/mapper/vg0-home}
\end{itemize-steps}

\subsubsection{Mount all the partitions}

\begin{itemize-steps}
	\item \terminalcmd{mount /dev/mapper/vg0-root /mnt}
	\item \terminalcmd{swapon /dev/mapper/vg0-swap}
	\item \terminalcmd{mkdir /mnt/boot}
	\item \terminalcmd{mount /dev/sda2 /mnt/boot}
	\item \terminalcmd{mkdir /mnt/boot/efi}
	\item \terminalcmd{mount /dev/sda1 /mnt/boot/efi}
	\item \terminalcmd{mkdir /mnt/home}
	\item \terminalcmd{mount /dev/mapper/vg0-home /mnt/home}
\end{itemize-steps}







\clearpage
\section{References}

\begin{itemize}
	\item \href{https://wiki.archlinux.org/index.php/MacBookPro11,x#Using_the_MacBook.27s_native_EFI_bootloader_.28recommended.29}{Arch Linux's MacBookPro 11.x Wiki page}
	\item \href{https://gist.github.com/mattiaslundberg/8620837}{Mattias Lindberg's Encrypted volume Arch install guide}
	\item \href{https://www.howtoforge.com/tutorial/how-to-install-arch-linux-with-full-disk-encryption/}{HowToForge | How to install Arch Linux with Full Disk Encryption}
\end{itemize}

\end{document}